---
title: "Class 11: Structural Bioinformatics 1"
author: "Pratik Varade"
date: "11/5/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## The PDB database for biomolecular structure data

> Q1: Determine the percentage of PDB structures solved by X-ray and Electron microscopy.

>What proportion of structures are protein. 

Download a CSV file from the PDB site (accessible from “Analyze” -> “PDB Statistics” > “by Experimental Method and Molecular Type”
```{r}
#read CSV file
data <- read.csv("Data Export Summary.csv")
data
View(data)
```
Total Number of Entries
```{r}
sum(data$Total)
```
Total number of entries from each method (XRAY NMR ELECTRON MICROSCOPY OTHER MULTIMETHOD respectively)
```{r}
(data$Total/sum(data$Total))*100
```
Proportion that are protein (round = round to how much ever sig figs)
```{r}
sum(data$Total)
sum(data$Proteins)

round ((sum(data$Proteins)/sum(data$Total)) *100, 2)
```



## HIV-Pr structure analysis
Can't see the hydrogens on VMD because the resolution of the experiement was too big to see the hydrogen molecules.

VMD > Graphics -> representation > and then you can have multiple representations at the same time by pressing Create Rep. Double click to enable and disable (red text).


Here we will read the 1HSG PDB structure and select the protein component and write out a new **protein - only** PDB formate file. We then do the same for the ligand (ie known drug molecule) creating a **ligand-only** PDB file.
```{r}
library(bio3d)
#read PDB file
pdb <- read.pdb("1hsg.pdb")
pdb

```

```{r}
ligand <- atom.select(pdb,"ligand", value= TRUE )
write.pdb(ligand, file="1hsg_ligand.pdb")
```


