---
title: "class08"
author: "Pratik Varade"
date: "10/24/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
Testing out how kmeans works.
```{r}
tmp <- c(rnorm(30,-3), rnorm(30,3))
x <- cbind(x=tmp, y=rev(tmp))
plot(x)
```


Use the kmeans() with k=2, nstart=20
```{r}
 k <- kmeans(x, centers =2, nstart=20)
```

Inspect/print results
```{r}
k
```
Q. How many points are in each cluster?
30 in each.

 Q. What ‘component’ of your result object details
 - cluster size?
 - cluster assignment/membership?
 - cluster center?


```{r}
k$size
k$cluster
k$centers
```

Plot x colored by the kmeans cluster assignment and
 add cluster centers as blue points

```{r}
plot(x, col = k$cluster)
points(k$centers, col="blue", pch=15)
```



Hierarchial Clustering in R


The 'hclust()' function requires a distance masteix as input.  You can get this from the 'dist()' function.

```{r}
# First we need to calculate point (dis)similarity
# as the Euclidean distance between observations
dist_matrix <- dist(x)

# The hclust() function returns a hierarchical
# clustering model
hc <- hclust(d = dist_matrix)

# the print method is not so useful here
hc
```
```{r}
plot(hc)
abline(h=6, col="red")
cutree(hc, h=6) # Cut by height h
```

```{r}
cutree(hc, k=2)
```


Linking Methods// A more real like example
```{r}
# Step 1. Generate some example data for clustering
x <- rbind(
 matrix(rnorm(100, mean=0, sd = 0.3), ncol = 2), # c1
 matrix(rnorm(100, mean = 1, sd = 0.3), ncol = 2), # c2
 matrix(c(rnorm(50, mean = 1, sd = 0.3), # c3
 rnorm(50, mean = 0, sd = 0.3)), ncol = 2))
colnames(x) <- c("x", "y")

# Step 2. Plot the data without clustering
plot(x)

# Step 3. Generate colors for known clusters
# (just so we can compare to hclust results)
col <- as.factor( rep(c("c1","c2","c3"), each=50) )
plot(x, col=col)
```





Q. Use the dist(), hclust(), plot() and cutree()
 functions to return 2 and 3 clusters 

```{r}

#clustering
h<- hclust(dist(x))

#draw tree
plot(h)
abline(h=2,col="red")

#Cut the tree into clusters/groups
grps <- cutree(h, k=3)
grps
```


Plot the data with grps colored
```{r}
plot(x, col=grps)
```





How many points in each cluster
```{r}
table(grps)
```


Cross= tablulate ie. compare out clustering result with the known answer
```{r}
table(grps, col)
```





Principal Component analysis (PCA)


```{r}
## You can also download this file from the class website!
mydata <- read.csv("https://tinyurl.com/expression-CSV",
 row.names=1)
head(mydata) 
```




How many genes are in this dataset?
```{r}
dim(mydata)
nrow(mydata)
```

Lets do PCA with the **prcomp** function.
```{r}
## lets do PCA
pca <- prcomp(t(mydata), scale=TRUE)
```
What are the attributes of pca?
```{r}
## See what is returned by the prcomp() function
attributes(pca)

```
Lets plot pca1 vs pca2
```{r}
## A basic PC1 vs PC2 2-D plot
plot(pca$x[,1], pca$x[,2]) 
```


```{r}
## Precent variance is often more informative to look at
pca.var <- pca$sdev^2
pca.var.per <- round(pca.var/sum(pca.var)*100, 1) 
pca.var.per
```
Make a scree plot with the pca.var.per data
```{r}
barplot(pca.var.per, main="Scree Plot",
 xlab="Principal Component", ylab="Percent Variation")
```
Color the points
```{r}
## A vector of colors for wt and ko samples
colvec <- colnames(mydata)
colvec[grep("wt", colvec)] <- "red"
colvec[grep("ko", colvec)] <- "blue"
plot(pca$x[,1], pca$x[,2], col=colvec, pch=16,
 xlab=paste0("PC1 (", pca.var.per[1], "%)"),
 ylab=paste0("PC2 (", pca.var.per[2], "%)")) 
```

```{r}
plot(pca$x[,1], pca$x[,2], col=colvec, pch=16,
 xlab=paste0("PC1 (", pca.var.per[1], "%)"),
 ylab=paste0("PC2 (", pca.var.per[2], "%)"))
## Click to identify which sample is which
identify(pca$x[,1], pca$x[,2], labels=colnames(mydata)) 

```


Hands On Activity
```{r}
x <- read.csv("UK_foods.csv", row.names = 1)
x
```
Q1. How many rows and columns are in your new data frame named x? What R functions could you use to answer this questions?
```{r}
dim(x) #OR
nrow(x)
ncol(x)

```
Checking My Data
```{r}
View(x)
```

Trying out different Plots
```{r}
barplot(as.matrix(x), beside=T, col=rainbow(nrow(x)))
```

Q3: Changing what optional argument in the above barplot() function results in the following plot? Change beside = T to beside= F
```{r}
barplot(as.matrix(x), beside=F, col=rainbow(nrow(x)))

```

```{r}
pairs(x, col=rainbow(10), pch=16)
```
Q6. What is the main differences between N. Ireland and the other countries of the UK in terms of this data-set?
Based on the plots above, N. Ireland is a lot more different than the other 3 counteries. It is not as linear as the rest. 

```{r}
# Use the prcomp() PCA function 
pca <- prcomp( t(x) )
summary(pca)
```

```{r}
plot (pca$x[,1], pca$x[,2])
```

























