---
title: "class14"
author: "Pratik Varade"
date: "11/12/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## Part 1 : Install Bioconductor/DESeq2 Setup
We will use the DESeq2 package from bioconductor. To install this we ran **in the console **
----**Note** eval= FALSE, just so it doesn't actually run because I already ran it in console.
```{r eval = FALSE}
install.packages("BiocManager")
BiocManager::install()
# For this class, you'll also need DESeq2:
BiocManager::install("DESeq2")
```

##Part 2:  Import countData and  colData
```{r}
counts <- read.csv("airway_scaledcounts.csv", stringsAsFactors = FALSE)
metadata <-  read.csv("airway_metadata.csv", stringsAsFactors = FALSE)
```

Take a look at these files :
```{r}
head(counts)
head(metadata)
```

How many genes do we have in this data set?
```{r}
nrow(counts)
```

How many exp do we have?
```{r}
ncol(counts)
```

Part 3: Toy differential gene expression
## perform some exploratory differential gene expression analysis
 
```{r}
View(metadata)
```
We have to find the mean of all the control groups data. We are doing this to find out if there is a difference in expression values for control (non-drug), vs treated (drug added to cell).



First step: Find all the control experiments, and then average all of them. Then we do the same for all the drug experiments.




```{r}
control <- metadata[metadata[,"dex"]=="control",]
control$id
```

Now we can calculate the mean across control columns of **countdata**

```{r}
control.mean <- rowSums( counts[ ,control$id] )/length(control$id)
names(control.mean) <- counts$ensgene
```

Now its time to do the same thing for the drug experiments.
```{r}
treated <- metadata[metadata[,"dex"]=="treated",]
treated$id
```
Now we can calc the mean across treated columns of countdata.
```{r}
treated.mean <- rowSums( counts[ ,treated$id] )/length(treated$id)
names(treated.mean) <- counts$ensgene
```


We will combine our meancount data for bookkeeping purposes.

```{r}
meancounts <- data.frame(control.mean, treated.mean)
```


Plot control vs Treated.
```{r}
plot(meancounts$control.mean, meancounts$treated.mean)
```


The plot above is too conjested in the beginning. I need to spread out the plot so trends can be seen.

```{r}
plot(meancounts$control.mean, meancounts$treated.mean, log="xy")
```


Make a log2 function.
```{r}
meancounts$log2fc <-log2(meancounts[,"treated.mean"]/meancounts[,"control.mean"])
head(meancounts)
```

The NaN and -inf is from doing math with 0's. We have to filter out the genes with 0 experession.
```{r}
zero.vals <- which(meancounts[,1:2]==0, arr.ind=TRUE)

to.rm <- unique(zero.vals[,1])
mycounts <- meancounts[-to.rm,]
head(mycounts)

```

**Testing what arr.ind = TRUE does**. It just tells you where on the dataset the 0 value is. So there are 3 0's (1, 2, 3), and the coordinates are given in the row/column. 
```{r}
x<- data.frame(control=c(56,0,5), treated=c(26,0,0))
x
which(x==0, arr.ind = TRUE)
```





```{r}
up.ind <- mycounts$log2fc > 2
down.ind <- mycounts$log2fc < (-2)
```
We sum up.ind because true = 1 and false = 0. So if we add them, we should get how many are True.
```{r}
sum(up.ind)
```

```{r}
head(mycounts[up.ind,])
```
Let’s do this the right way. DESeq2 is an R package for analyzing count-based NGS data like RNA-seq. It is available from Bioconductor. Bioconductor is a project to provide tools for analyzing high-throughput genomic data including RNA-seq, ChIP-seq and arrays.

```{r}
library(DESeq2)
citation("DESeq2")
```

We will use the DESeqDataSetFromMatrix() function to build the required DESeqDataSet object and call it dds, short for our DESeqDataSet. 

```{r}
dds <- DESeqDataSetFromMatrix(countData=counts, 
                              colData=metadata, 
                              design=~dex, 
                              tidy=TRUE)
dds
```


Run DESEQ2
```{r}
dds <- DESeq(dds)
```
Look at the results
```{r}
res <- results(dds)
res
```

## Making a summary plot (Volcano Plot)
This figure will combine both Change and the p-value into the overview figure indicating the proportion of genes with large scale significant difference in their expression.

```{r}
#This graph is dull, and hard to read. Add color.
plot(res$log2FoldChange, -log(res$padj))
abline(v=c(-2,2), col="gray", lty=2)
abline(h=-log(0.05), col="gray", lty = 2)
```

```{r}
mycol <- rep("gray", nrow(res))
#make points with +2 -2 fold change blue
mycol [abs(res$log2FoldChange) >2] <- "blue"
#make points below pvalue cutoff red.
plot( res$log2FoldChange,  -log(res$padj), 
 col=res$sig+1, ylab="-Log(P-value)", xlab="Log2(FoldChange)")

# Add some cut-off lines
abline(v=c(-2,2), col="darkgray", lty=2)
abline(h=-log(0.1), col="darkgray", lty=2)
```
Set up some custom color so we can read the plot better.
```{r}
# Setup our custom point color vector 
mycols <- rep("gray", nrow(res))
mycols[ abs(res$log2FoldChange) > 2 ]  <- "red" 

inds <- (res$padj < 0.01) & (abs(res$log2FoldChange) > 2 )
mycols[ inds ] <- "blue"

#Volcano plot with custom colors 
plot( res$log2FoldChange,  -log(res$padj), 
 col=mycols, ylab="-Log(P-value)", xlab="Log2(FoldChange)" )

abline(v=c(-2,2), col="gray", lty=2)
abline(h=-log(0.1), col="gray", lty=2)
```

Save our results for the next day
```{r}
write.csv(res, file="expression_results.csv")
```







































