---
title: "Class 7 R functions and packages"
author: "Pratik Varade"
date: "10/22/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R functions revisited

Source my functions from last day
```{r}
source("http://tinyurl.com/rescale-R")
```

```{r}
rescale(1:10)
```

```{r}
rescale(c(1,10,5,NA,6))
```
```{r, eval=FALSE}
rescale2(c(1:10,"barry"))
```

```{r , eval= FALSE}
rescale(c(1:10, "Barry"))
```
The error on the rescale 2 function is a lot more helpful than rescale.

##Write a function to find NA in 2 vectors
Write a function to find where there are NA elements in two input vectors.
First make some simple input where I know the answers.
```{r}
x <- c( 1, 2, NA, 3, NA)
y <- c(NA, 3, NA, 3, 4)
```
Looked online and found the **is.na()** function
```{r}
is.na(x)
```
and the **which()** function tells me where the TRUE values are
```{r}
which(is.na(x))
```

```{r}
which(is.na(y))
```

```{r}
is.na(x)
is.na(y)
```
the AND function requires two input TRUE to give a TRUE output
```{r}
is.na(x) & is.na(y)
```
Taking the **sum()** of TRUE FALSE vector will tell me how many TRUE elements I have. Remember, TRUE = 1, FALSE = 0. So when sum = 1, that means there is one pair of TRUES. This is my working **snippet**.
```{r}
sum(is.na(x)& is.na(y))
```

```{r}
sum(c(T, T, F, T))
```
Now turn that **snippet** into a function.

```{r}
both_NA <- function(x, y) {
sum(is.na(x) & is.na(y))
}
```

Test
```{r}
both_NA(x,y)
```
Eejit Proofing, both_NA(x,y2) is weird because one vector has 3 elements, while the other has 4. What happens is recycling. x is too small, so it will recycle it, so in theory its NA, NA, NA, (original 3 NAs) NA, NA, NA(this is recycled x vector), but only the first 4 count because the second vector is 4 elements.
```{r}
x <- c(NA, NA, NA)
y1 <- c( 1, NA, NA)
y2 <- c( 1, NA, NA, NA)

both_NA(x,y1)
both_NA(x,y2)
```

```{r}
x2<- c(NA,NA)
both_NA(x2,y2)
```

Testing how recycling works.
```{r}
x <-   c(1, NA, NA)
y3 <- c( 1, NA, NA, NA, NA, NA, NA)

both_NA(x,y3)
```


This is the same as the problem on top because of recycling.
```{r}
x3<- c(1, NA,NA,1,NA,NA,1)
y3 <- c(1, NA, NA, NA,NA,NA, NA)
```

```{r}
length(x)
length(y3)
```


Add a check for when inputs x and u are not the same lengths
```{r}
both_NA2 <- function(x, y) {
if(length(x) != length(y)) {
stop(print("Inputs x and y should be the same length!"))
  }
  sum(is.na(x) & is.na(y))
}

```


Test
```{r eval= FALSE}
both_NA2(x,y3)
```


Make a function that will help Prof. Barry grade student work.
```{r}
# student 1
student1 <- c(100, 100, 100, 100, 100, 100, 100, 90)
# student 2
student2 <- c(100, NA, 90, 90, 90, 90, 97, 80)
```

Student 1
```{r}
mean(student1[-which.min(student1)])
```

Student 2
```{r}
which.min(student2)
student2[-which.min(student2)]
mean(student2[-which.min(student2)], na.rm= TRUE)
```




any means if any of the vector is NA, it will say true. If all of them are not NA, false.
```{r}
is.na(student2)
any(is.na(student2))
any(is.na(student1))
```
We have out working code now turn it into a first function

```{r}
grade1 <- function(x) {
  if(any(is.na(x)))
    warning(print("This student is missing a homework"))
  mean(x[-which.min(x)], na.rm=TRUE)
}
```



We have out working code now turn it into a first function
```{r, eval = FALSE}
grade2 <- function(x) {
  if(any(is.na(x)))
    warning(print("This student is missing a homework"))
  sum(x[-which.min(x)], na.rm=TRUE)/length(x)-1
}

grade(student1)
grade(student2)
```

```{r}
s3<- c(100, NA,NA,NA,NA)
```


```{r}
url <- "https://tinyurl.com/gradeinput"
read.csv(url)
hw <- read.csv(url, row.names=1)
```

1 = rows, 2 = columns
```{r}
apply(hw, 1, grade1)
```


Datapasta
```{r , eval= FALSE}
install.packages("datapasta")

```


























